The Talking Heads Experiment

### 2015-04-10 chap11/chap11.tex
fixes
quotation marks for chapters 8 to 10

While introducing these experiments I will highlight two conceptual breakthroughs: semiotic networks and language strategies. 
\begin{itemize}
\item A {\itshape semiotic network} consists of a weighted network that links words to concepts, concepts to prototypes, and prototypes to 
sensory experiences. Each agent constructs his own semiotic network but through consecutive 
interactions these networks become progressively similar. From this perspective, setting up a language game requires a definition and 
operationalisation how new nodes are created in this network, how new links are made between 
the nodes, and how the weights between nodes are adjusted. 
\item A {\itshape language strategy} is helpful to think about how we can go beyond individual words and towards 
more complex grammatical expressions. It was already introduced in the previous chapter. 
A language strategy is a particular way to deal syntactically and semantically 
with one domain of meaning. It includes ways for conceptualising, expressing, acquiring and expanding a specific domain. 
For example, the colour domain requires not only strategies for basic colour terms (like ``blue"), but also for 
graded membership expressions (``very blue"), composite colours, (``greenish blue"), analogical colours (``sky blue"), a.o. 
The Talking Heads experiments used a single strategy that was entirely hard-coded: discrimination trees for categorisation 
and words without syntax to express distinctive categories. Research efforts moved towards the study of other strategies, and 
to the question how such strategies may arise (discussed in the next chapter). 
\end{itemize}
The remainder of the present chapter illustrates these two notions through different language game experiments: The Proper 
Naming Game and the Action Naming Game illustrate the use of semiotic networks. The Colour Description Game and the 
Spatial Language Game illustrate the use of language strategies. These experiments also show how the focus of research 
progressively moved towards the emergence of grammar in co-evolution with more complex compositional, but still 
grounded, semantics.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap11/figs/grounded-naming-game}}
\caption{\label{fig:grounded} 
Set-up for the Grounded Proper Naming Game. Two Qrio robots are playing a game of reference about individual objects 
in their environment. They are using proper names and hence have to recognise objects as individuals.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap10/figs/png-perception}}
\caption{\label{fig:png-perception} 
A robot is looking at consecutive scenes (from top to bottom) and tracking the objects in the scene. One object (labeled $O_{716}$)
is moved by a human experimenter.  We see from left to right: the original source image, results of region detection and 
colour segmentation, the (average) value for one dimension of one object ($O_{716}$), the segmented and labeled objects, 
and the world model, including the position of the other robot (indicated by a pointing arrow). }
\end{figure}

@@

We have already seen from earlier experiments that the solution to these various issues lies in setting up
the right kind of `semiotic dynamics'. In the present case, this dynamics should gradually coordinate
sensory experiences, prototypical views, individuals, 
and names, both within a single individual and across the
population. This lead to the notion of a {\itshape semiotic network}, which was a significant conceptual advance 
in the methodology of conceiving and carrying out language game experiments.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap10/figs/prototypes-identifiers-words}}
\caption{\label{fig:prototypes-identifiers-words} 
In the Proper Naming Game experiment the memory of the agents is conceived in terms of a semiotic network that 
spans the full semiotic cycle from sensory experience and conceptualisation to naming.}
\end{figure}

@@

Given this notion of a network, the key question for designing an effective 
strategy for playing the language game becomes: What are the operators that are building and
changing the semiotic networks in each agent? We focus first on the
vocabulary, i.e. the links between nodes for individual objects $m_i$ and their 
names $f_i$. We use a variant of the strategy used in the Talking Heads experiment. 
Only the hearer changes the score after a game. 
In the case of a successful game, the score $\sigma_{f_i,m_i}$ of the used association is 
increased and its competitors are decreased according to the following
equations with the alignment rate $\gamma = 0.2$: 
\begin{equation}
 \sigma_{m_i,f_i} \leftarrow \sigma_{m_i,f_i} (1-\gamma) + \gamma
\end{equation}
\begin{equation}
 \sigma_{m_j,f_i} \leftarrow \sigma_{m_j,f_i} (1-\gamma) ~~j \neq i
\end{equation}
\begin{equation}
 \sigma_{m_i,f_k} \leftarrow \sigma_{m_i,f_k} (1-\gamma) ~~k \neq i
\end{equation}
A competitor is another individual $m_j$ stored in the hearer's memory 
for the name $f_i$ or another name $f_k$ for the individual 
$m_i$. When the speaker is using a name which is not linked in the network 
of the hearer to any individual, then the hearer adds a new relation between 
the recognised individual (after pointing by the speaker) and this new name. 
This alternative then competes from 
now on with existing associations through the lateral inhibition dynamics. 
We already know from many other experiments that this strategy, when used collectively in 
consecutive games between randomly chosen members of the population, leads to the  
self-organisation of a shared lexicon due to positive feedback.\footnote{
The semiotic dynamics of lateral inhibition has been studied thoroughly by Bart de Vylder as reported in his Ph.D thesis 
(2007 at the VUB AI Lab) and summarized in this paper: \cite{deVylder:2006}.}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap11/figs/png-results}}
\caption{\label{fig:png-results} 
{\itshape Left:} Results for Proper Naming games in a population of 10 agents that have to name 15 individual objects. 
We see that a vocabulary establishes
itself after agents have played a few hundred games. There is again the typical peak until alignment kicks in. 
The number of individuals is larger than 15, implying that agents name prototypical views rather than 
individuals. {\itshape Right:} Results after the addition  
of a heuristic based on tracking an object over time. We see that there are fewer individuals 
than stored prototypical views, which means that agents have improved their ability to recognise object 
identity. }
\end{figure}
With these mechanisms, the population reaches a high level of communicative success (above 90 \%). 
However the average number of individuals in the agents' semiotic networks
is much larger than the number of distinctive objects
introduced in the experiment (see \figref{fig:png-results}, left). Apparently, agents are naming prototypical views of
individual objects instead of the individuals themselves. So we did not adequately address 
how a robot learns to interrelate different views. To do this, various heuristics need to be added.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.75\textwidth]{chap11/figs/adjust-id}}
\caption{\label{fig:png-optimisation} 
When the identity of two objects is discovered (because two prototypical views have been established 
as belonging to the same individual object), the former nodes for identities can be merged while keeping the 
links to their names intact. 
}
\end{figure}

@@

Clearly humans use many additional heuristics. For example, if we see somebody walking into a
building with a refrigerator and we later see the same person on the
top floor handling a refrigerator we assume it is the same
refrigerator, even if we could not track this object. Or if we know that a particular person is 
going to come and visit, and next we hear the bell and open the door, we expect and recognise this particular 
person even if clothes or hair have changed drastically. The
point of the present experiment was not to operationalise all imaginable heuristics but to
show that heuristics help to optimize and coordinate semiotic
networks and thus further increase the ability of a population to develop internal shared
representations anchored in the world.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/myon-action-game}}
\caption{
Two Myon robots playing an action game. One robot (the speaker) names an action and the other robot (the hearer) 
has to perform the action. When the hearer performs the right action the speaker signals success by nodding the head, 
otherwise he performs the action himself so that the hearer can adjust his semiotic network accordingly. {\itshape Left:} the speaker
requests an action. {\itshape Middle:} the hearer interprets this as moving up both arms. {\itshape Right:} correction by the speaker. 
}\label{fig:myon-action} 
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.2\linewidth]{chap11/figs/postures.pdf}\includegraphics[width=0.75\linewidth]{chap11/figs/vision_2.pdf}}
\caption{\label{vision} \footnotesize{{\itshape Left:} Example postures of the robot (which is the QRIO in this case).
{\itshape Right:} Visual analysis. From left to right we see the source image, the foreground/background distinction and object segmentation (focusing on the upper torso), and the feature signature of this posture represented as a graph connecting the values (on the y-axis)
of the seven centralized normalized moments (on the x-axis).}}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.6\linewidth]{chap11/figs/robot-mirror}}
\caption{\label{robot-mirror} \footnotesize{The QRIO humanoid robot stands before a mirror and performs various motor behaviours thus observing what visual body-images these behaviours generate.}}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.6\linewidth]{chap11/figs/commsuccess}}
\caption{\label{commsuccess} \footnotesize{Results of the Action Game played by a population of 10 agents 
for 10 postures after coordinating image schemata and motor control through a mirror. 
The x-axis plots the number of language games. The y-axis shows the running average of 
communicative success and average lexicon size.}}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.6\linewidth]{chap11/figs/network}}
\caption{ \label{network} Semiotic network linking sensory experiences, visual prototypes of postures, nodes for postures acting as mirror neurons, and nodes triggering the motor behaviour that achieves the posture. Nodes for words (shown as squares) 
are associated with the posture nodes.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap11/figs/colour-qrio}}
\caption{\label{fig:munsell} 
Set-up with two QRIO humanoids, similar to Munsell chip tests used in psychology. The same game as the Talking 
Heads was used, but only colour chips were provided as input and the robots use pointing and head movements to 
provide feedback.}
\end{figure}

When the QRIO humanoid robots became available, we decided to use this existing work for pushing forward the state of 
the art in colour language research in two ways:
\begin{itemize}
\item To use more realistic circumstances of object perception. Because 
robots are moving around in three dimensions and see the situation from different points of view, the reflection 
of light changes and is different for speaker and hearer (see \tabref{f:scene-example}). This difference did not 
play a role in earlier experiments but 
it certainly plays an important role in the stabilisation and choice of colour prototypes. 
\item To focus on colour descriptions beyond basic colour terms, i.e. descriptions like 
``slightly blue", ``blue-green", ``bluish green", ``very bright" ``shiny yellow", ``dark red", ``sky blue", 
``very dark bluish red", etc. Indeed, it turns out that 
basic colour terms are used only in about 10 to 15 \% of the occurrences of colour descriptions. The rest are 
more complex expressions. Complex colour descriptions require scaling up the semantics and introducing grammar, they 
also require the introduction of language strategies as a layer of abstraction 
above lexical and grammatical constructions. 
\end{itemize}

@@

    &\includegraphics[width=0.37\textwidth]{chap11/figs/grounding-scene-a}
	&\includegraphics[width=0.37\textwidth]{chap11/figs/grounding-scene-b}\\
  \multicolumn{2}{c}{
    \begin{tabular}{lrrr}
	      & obj-9 & obj-11 & obj-17 \\
    \midrule
    \emph{L*} &  35.5 &  51.2 & 50.5 \\
    \emph{a*} &   7.7 & -17.1 & 26.7 \\
    \emph{b*} & -40.7 & -14.0 & 39.6 \\
    \end{tabular}
  }
&
  \begin{tabular}{rrr}
 obj-7 & obj-10 & obj-15 \\
\midrule
      35.6 & 52.8 & 62.2 \\
      7.2 & -20.1 & 27.9 \\
      -39.0 & -11.3 & 52.5 \\ 
  \end{tabular} \\
\lspbottomrule
\end{tabular}
  \caption[Comparison between the colour perceptions of two robots for
  an example scene]{Colour perceptions of two
    robots for the same scene. The robots see the yellow duck
    (obj-17 for the left robot and obj-15 for the right robot) from
    different sides and distances and thus perceive very different colour values on 
    the \emph{a*} and \emph{b*} colour dimensions for the same object.}
  \label{f:scene-example}
\end{table}

@@

A language strategy has two components: (i) a particular way of conceptualising reality (and learning the categories involved
in this conceptualisation) and (ii) a particular grammatical pattern that suggests to the listener what strategy is intended 
and provides the contents to execute the strategy. For example, in the phrase ``very green", the combination of an adverb 
``very" and a basic colour adjective ``green" suggests a `graded membership strategy' where not only a basic colour prototype
('green') is introduced but also how close the sample is to this prototype ('very'). I will now introduce first how 
the conceptual aspect of a language strategy is defined using a procedural semantics system called IRL and then how the 
linguistic aspect is defined, using Fluid Construction Grammar. \\

@@

Constraint networks are an obvious choice for grounded (natural) language processing. \is{constraint network}
For example, to interpret the sentence ``the ball behind
the block", it is possible that there are many blocks, many balls, but only one ball that is behind a block. So the 
unique referent of ``the ball", ``the block" and of the total noun phrase only become clear when 
all constraints have been applied. Natural languages clearly do not supply control information but assume that 
the interpreter is smart enough to interpret constraints based on data flow. The other properties of constraint 
programming are also important. The constraint network planner can be used for conceptualising what to say and 
the constraint network matcher is highly useful in parsing an incomplete language fragment or in learning the 
meaning of a new grammatical construction. \\

@@

What is the nature of the constraints that we will need for the semantics of natural language? 
Take a phrase like ``the ball behind the block". It invokes the following constraints, also called 
semantic operations: 
\begin{itemize}
\item ``the ball" evokes a constraint to use a prototype [BALL] to find a referent in the current scene.
\item The same holds for ``the block" which uses another prototype [BLOCK]. 
\item The preposition ``behind" implies another constraint, namely to use the spatial relation [BEHIND] to filter a 
set of objects (namely those found by using the [BALL] prototype) and retain only those that satisfy the behind-relation 
to an object in another set. 
\end{itemize}

@@

It should be possible for the filter-using-prototype operation
to either apply a filler for ?prototype-1 and 
?context and then compute ?object-set-1, or provide ?object-set-1 and ?context and obtain a filler for ?prototype-1. 
The 'normal' use is the first one, but the second use is just as important in planning what to
say (as speaker) or in guessing a possible prototype while learning the meaning of some unknown word. 
Moreover the `identify-set-using-prototype' operation can be made smart enough that it can actually come up with
a new prototype, when there isn't one yet that does the job. Hence learning can be 
integrated directly into the constraint network itself.

@@

The constraint network planner works on an evolutionary basis, similar to genetic algorithms. Starting from the goal 
(for example identifying a particular object in the environment), it will generate possible networks by recruiting 
existing primitive operations, linking them in, and checking whether the computation makes progress towards reaching the goal. 
The slots of a constraint are typed to limit what semantic operations will be tried next. Networks that perform useful 
functions are `chunked' into higher order units that then can be used as units in their own right, thus scaling up 
complexity. These chunked networks become associated with stereotyped grammatical expression, so that there is a very 
fast retrieval and semantic execution. \\

@@

1. The first strategy is the {\bfshape Basic Colour Strategy} which implements a prototype-based approach to 
the basic colour categories. This approach is similar to the one used in the Proper Naming Game, except that 
only the three colour-dimensions are profiled (the opponent channels red-green, yellow-blue and the brightness channel). 
The strategy is similar to the one shown earlier. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour-lenient ?s3 ?s2 ?cc)
(select-most-activated ?t ?s3)
\end{verbatim}
When the speaker says ``green", this network is used together with a specific binding of ?cc to the prototype for 
green (denoted as [green]). 
\begin{verbatim}
(bind colour-category ?cc [green])
\end{verbatim}
A graphical representation of this network is shown in \figref{fig:basic-strat}.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/basic-strat}}
\caption{\label{fig:basic-strat} 
Constraint network for the Basic Colour strategy. The colours of each object in the context are retrieved, their 
similarity to the prototype computed, and the one with the highest similarity is chosen.}
\end{figure}

2. The second strategy is the {\bfshape Graded Membership Strategy} which categorises the topic first using 
a basic colour but then qualifies how well its colour fits with the prototype. For example, we can say ``greenish" or
``very green". Some languages (such as Tarahumara spoken in Northern Mexico) obligatory mark graded membership using 
suffixes. The core of the constraint network is defined as follows. A graphical representation of the network is shown
in \figref{fig:graded}. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour ?s3 ?s2 ?cc)
(filter-by-membership ?s4 ?s3 ?mc)
(select-most-activated ?t ?s4)
\end{verbatim}
This network grabs the objects in the context and binds them to ?s1 using the equal-to-context operation. 
The profile-colour-dimensions operation picks out the colour properties of the various objects in ?s1 resulting in 
a new set ?s2. The filter-by-colour operation computes for each object in the set ?s2
how well they match with the prototype bound to ?cc, constructing a new set ?s3. The filter-by-membership operation 
then computes how well 
each object in ?s3 satisfies the graded membership qualification bound to ?mc, 
constructing ?s4. And then the select-most-activated operation picks out the best candidate and binds it to ?t. 
Other data flows within the same network are possible, for example to find the bindings of ?mc and ?cc given 
a context and a referent ?t that needs to be identified.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/graded}}
\caption{\label{fig:graded} 
Constraint network implementing the Graded Membership Strategy. It builds further on the network for Basic Colour
Categorisation but adds a constraint to handle graded membership. }
\end{figure}

The network is entirely general for all expressions that use graded membership and gets instantiated for a particular 
case by specifying the semantic entities involved using the bind operation. For example, Tarahumara has the word 
``sita" for a particular colour prototype and ``kame" for one type of graded membership: 
\begin{verbatim}
(bind colour-category ?cc [sita])
(bind membership-category ?mc [kame]) 
\end{verbatim}

>>>>

While introducing these experiments I will highlight two conceptual breakthroughs: semiotic networks and language strategies. 
\begin{itemize}
\item A {\itshape semiotic network} consists of a weighted network that links words to concepts, concepts to prototypes, and prototypes to 
sensory experiences. Each agent constructs his own semiotic network but through consecutive 
interactions these networks become progressively similar. From this perspective, setting up a language game requires a definition and 
operationalisation how new nodes are created in this network, how new links are made between 
the nodes, and how the weights between nodes are adjusted. 
\item A {\itshape language strategy} is helpful to think about how we can go beyond individual words and towards 
more complex grammatical expressions. It was already introduced in the previous chapter. 
A language strategy is a particular way to deal syntactically and semantically 
with one domain of meaning. It includes ways for conceptualising, expressing, acquiring and expanding a specific domain. 
For example, the colour domain requires not only strategies for basic colour terms (like `blue'), but also for 
graded membership expressions (`very blue'), composite colours, (`greenish blue'), analogical colours (`sky blue'), a.o. 
The Talking Heads experiments used a single strategy that was entirely hard-coded: discrimination trees for categorisation 
and words without syntax to express distinctive categories. Research efforts moved towards the study of other strategies, and 
to the question how such strategies may arise (discussed in the next chapter). 
\end{itemize}
The remainder of the present chapter illustrates these two notions through different language game experiments: The Proper 
Naming Game and the Action Naming Game illustrate the use of semiotic networks. The Colour Description Game and the 
Spatial Language Game illustrate the use of language strategies. These experiments also show how the focus of research 
progressively moved towards the emergence of grammar in co-evolution with more complex compositional, but still 
grounded, semantics.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap11/figs/grounded-naming-game.pdf}}
\caption{\label{fig:grounded} 
Set-up for the Grounded Proper Naming Game. Two Qrio robots are playing a game of reference about individual objects 
in their environment. They are using proper names and hence have to recognise objects as individuals.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap10/figs/png-perception.pdf}}
\caption{\label{fig:png-perception} 
A robot is looking at consecutive scenes (from top to bottom) and tracking the objects in the scene. One object (labeled $O_{716}$)
is moved by a human experimenter.  We see from left to right: the original source image, results of region detection and 
colour segmentation, the (average) value for one dimension of one object ($O_{716}$), the segmented and labeled objects, 
and the world model, including the position of the other robot (indicated by a pointing arrow). }
\end{figure}

@@

We have already seen from earlier experiments that the solution to these various issues lies in setting up
the right kind of ``semiotic dynamics''. In the present case, this dynamics should gradually coordinate
sensory experiences, prototypical views, individuals, 
and names, both within a single individual and across the
population. This lead to the notion of a {\itshape semiotic network}, which was a significant conceptual advance 
in the methodology of conceiving and carrying out language game experiments.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap10/figs/prototypes-identifiers-words.pdf}}
\caption{\label{fig:prototypes-identifiers-words} 
In the Proper Naming Game experiment the memory of the agents is conceived in terms of a semiotic network that 
spans the full semiotic cycle from sensory experience and conceptualisation to naming.}
\end{figure}

@@

Given this notion of a network, the key question for designing an effective 
strategy for playing the language game becomes: What are the operators that are building and
changing the semiotic networks in each agent? We focus first on the
vocabulary, i.e. the links between nodes for individual objects $m_i$ and their 
names $f_i$. We use a variant of the strategy used in the Talking Heads experiment. 
Only the hearer changes the score after a game. 
In the case of a successful game, the score $\sigma_{f_i,m_i}$ of the used association is 
increased and its competitors are decreased according to the following
equations with the alignment rate $\gamma = 0.2$: 
\begin{equation}
 \sigma_{m_i,f_i} \leftarrow \sigma_{m_i,f_i} (1-\gamma) + \gamma
\end{equation}
\begin{equation}
 \sigma_{m_j,f_i} \leftarrow \sigma_{m_j,f_i} (1-\gamma) ~~j \neq i
\end{equation}
\begin{equation}
 \sigma_{m_i,f_k} \leftarrow \sigma_{m_i,f_k} (1-\gamma) ~~k \neq i
\end{equation}
A competitor is another individual $m_j$ stored in the hearer's memory 
for the name $f_i$ or another name $f_k$ for the individual 
$m_i$. When the speaker is using a name which is not linked in the network 
of the hearer to any individual, then the hearer adds a new relation between 
the recognised individual (after pointing by the speaker) and this new name. 
This alternative then competes from 
now on with existing associations through the lateral inhibition dynamics. 
We already know from many other experiments that this strategy, when used collectively in 
consecutive games between randomly chosen members of the population, leads to the  
self-organisation of a shared lexicon due to positive feedback.\footnote{
The semiotic dynamics of lateral inhibition has been studied thoroughly by Bart de Vylder as reported in his Ph.D thesis 
(2007 at the VUB AI Lab) and summarised in this paper: \cite{deVylder:2006}.}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap11/figs/png-results.pdf}}
\caption{\label{fig:png-results} 
{\itshape Left:} Results for Proper Naming games in a population of 10 agents that have to name 15 individual objects. 
We see that a vocabulary establishes
itself after agents have played a few hundred games. There is again the typical peak until alignment kicks in. 
The number of individuals is larger than 15, implying that agents name prototypical views rather than 
individuals. {\itshape Right:} Results after the addition  
of a heuristic based on tracking an object over time. We see that there are fewer individuals 
than stored prototypical views, which means that agents have improved their ability to recognise object 
identity. }
\end{figure}
With these mechanisms, the population reaches a high level of communicative success (above 90 \%). 
However the average number of individuals in the agents' semiotic networks
is much larger than the number of distinctive objects
introduced in the experiment (see \figref{fig:png-results}, left). Apparently, agents are naming prototypical views of
individual objects instead of the individuals themselves. So we did not adequately address 
how a robot learns to interrelate different views. To do this, various heuristics need to be added.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.75\textwidth]{chap11/figs/adjust-id.pdf}}
\caption{\label{fig:png-optimisation} 
When the identity of two objects is discovered (because two prototypical views have been established 
as belonging to the same individual object), the former nodes for identities can be merged while keeping the 
links to their names intact. 
}
\end{figure}

@@

Clearly humans use many additional heuristics. For example, if we see somebody walking into a
building with a refrigerator and we later see the same person on the
top floor handling a refrigerator we assume it is the same
refrigerator, even if we could not track this object. Or if we know that a particular person is 
going to come and visit, and next we hear the bell and open the door, we expect and recognise this particular 
person even if clothes or hair have changed drastically. The
point of the present experiment was not to operationalise all imaginable heuristics but to
show that heuristics help to optimise and coordinate semiotic
networks and thus further increase the ability of a population to develop internal shared
representations anchored in the world.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/myon-action-game.pdf}}
\caption{
Two Myon robots playing an action game. One robot (the speaker) names an action and the other robot (the hearer) 
has to perform the action. When the hearer performs the right action the speaker signals success by nodding the head, 
otherwise he performs the action himself so that the hearer can adjust his semiotic network accordingly. {\itshape Left:} the speaker
requests an action. {\itshape Middle:} the hearer interprets this as moving up both arms. {\itshape Right:} correction by the speaker. 
}\label{fig:myon-action} 
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.2\linewidth]{chap11/figs/postures.pdf}\includegraphics[width=0.75\linewidth]{chap11/figs/vision_2.pdf}}
\caption{\label{vision} \footnotesize{{\itshape Left:} Example postures of the robot (which is the QRIO in this case).
{\itshape Right:} Visual analysis. From left to right we see the source image, the foreground/background distinction and object segmentation (focusing on the upper torso), and the feature signature of this posture represented as a graph connecting the values (on the y-axis)
of the seven centralised normalised moments (on the x-axis).}}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.6\linewidth]{chap11/figs/robot-mirror.pdf}}
\caption{\label{robot-mirror} \footnotesize{The QRIO humanoid robot stands before a mirror and performs various motor behaviours thus observing what visual body-images these behaviours generate.}}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.6\linewidth]{chap11/figs/commsuccess.pdf}}
\caption{\label{commsuccess} \footnotesize{Results of the Action Game played by a population of 10 agents 
for 10 postures after coordinating image schemata and motor control through a mirror. 
The x-axis plots the number of language games. The y-axis shows the running average of 
communicative success and average lexicon size.}}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.6\linewidth]{chap11/figs/network.pdf}}
\caption{ \label{network} Semiotic network linking sensory experiences, visual prototypes of postures, nodes for postures acting as mirror neurons, and nodes triggering the motor behaviour that achieves the posture. Nodes for words (shown as squares) 
are associated with the posture nodes.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap11/figs/color-qrio.pdf}}
\caption{\label{fig:munsell} 
Set-up with two QRIO humanoids, similar to Munsell chip tests used in psychology. The same game as the Talking 
Heads was used, but only colour chips were provided as input and the robots use pointing and head movements to 
provide feedback.}
\end{figure}

When the QRIO humanoid robots became available, we decided to use this existing work for pushing forward the state of 
the art in colour language research in two ways:
\begin{itemize}
\item To use more realistic circumstances of object perception. Because 
robots are moving around in three dimensions and see the situation from different points of view, the reflection 
of light changes and is different for speaker and hearer (see \tabref{f:scene-example}). This difference did not 
play a role in earlier experiments but 
it certainly plays an important role in the stabilisation and choice of colour prototypes. 
\item To focus on colour descriptions beyond basic colour terms, i.e. descriptions like 
`slightly blue', `blue-green', `bluish green', `very bright', `shiny yellow', `dark red', `sky blue', 
`very dark bluish red', etc. Indeed, it turns out that 
basic colour terms are used only in about 10 to 15 \% of the occurrences of colour descriptions. The rest are 
more complex expressions. Complex colour descriptions require scaling up the semantics and introducing grammar, they 
also require the introduction of language strategies as a layer of abstraction 
above lexical and grammatical constructions. 
\end{itemize}

@@

    &\includegraphics[width=0.37\textwidth]{chap11/figs/grounding-scene-a.pdf}
	&\includegraphics[width=0.37\textwidth]{chap11/figs/grounding-scene-b.pdf}\\
  \multicolumn{2}{c}{
    \begin{tabular}{lrrr}
	      & obj-9 & obj-11 & obj-17 \\
    \midrule
    \emph{L*} &  35.5 &  51.2 & 50.5 \\
    \emph{a*} &   7.7 & -17.1 & 26.7 \\
    \emph{b*} & -40.7 & -14.0 & 39.6 \\
    \end{tabular}
  }
&
  \begin{tabular}{rrr}
 obj-7 & obj-10 & obj-15 \\
\midrule
      35.6 & 52.8 & 62.2 \\
      7.2 & -20.1 & 27.9 \\
      -39.0 & -11.3 & 52.5 \\ 
  \end{tabular} \\
\lspbottomrule
\end{tabular}
  \caption[Comparison between the colour perceptions of two robots for
  an example scene]{Colour perceptions of two
    robots for the same scene. The robots see the yellow duck
    (obj-17 for the left robot and obj-15 for the right robot) from
    different sides and distances and thus perceive very different colour values on 
    the \emph{a*} and \emph{b*} colour dimensions for the same object.}
  \label{f:scene-example}
\end{table}

@@

A language strategy has two components: (i) a particular way of conceptualising reality (and learning the categories involved
in this conceptualisation) and (ii) a particular grammatical pattern that suggests to the listener what strategy is intended 
and provides the contents to execute the strategy. For example, in the phrase `very green', the combination of an adverb 
`very' and a basic colour adjective `green' suggests a ``graded membership strategy'' where not only a basic colour prototype
(`green') is introduced but also how close the sample is to this prototype (`very'). I will now introduce first how 
the conceptual aspect of a language strategy is defined using a procedural semantics system called IRL and then how the 
linguistic aspect is defined, using Fluid Construction Grammar. \\

@@

Constraint networks are an obvious choice for grounded (natural) language processing. \is{constraint network}
For example, to interpret the sentence `the ball behind
the block', it is possible that there are many blocks, many balls, but only one ball that is behind a block. So the 
unique referent of `the ball', `the block' and of the total noun phrase only become clear when 
all constraints have been applied. Natural languages clearly do not supply control information but assume that 
the interpreter is smart enough to interpret constraints based on data flow. The other properties of constraint 
programming are also important. The constraint network planner can be used for conceptualising what to say and 
the constraint network matcher is highly useful in parsing an incomplete language fragment or in learning the 
meaning of a new grammatical construction. \\

@@

What is the nature of the constraints that we will need for the semantics of natural language? 
Take a phrase like `the ball behind the block'. It invokes the following constraints, also called 
semantic operations: 
\begin{itemize}
\item `the ball' evokes a constraint to use a prototype [BALL] to find a referent in the current scene.
\item The same holds for `the block' which uses another prototype [BLOCK]. 
\item The preposition `behind' implies another constraint, namely to use the spatial relation [BEHIND] to filter a 
set of objects (namely those found by using the [BALL] prototype) and retain only those that satisfy the behind-relation 
to an object in another set. 
\end{itemize}

@@

It should be possible for the filter-using-prototype operation
to either apply a filler for ?prototype-1 and 
?context and then compute ?object-set-1, or provide ?object-set-1 and ?context and obtain a filler for ?prototype-1. 
The 'normal' use is the first one, but the second use is just as important in planning what to
say (as speaker) or in guessing a possible prototype while learning the meaning of some unknown word. 
Moreover the ``identify-set-using-prototype'' operation can be made smart enough that it can actually come up with
a new prototype, when there isn't one yet that does the job. Hence learning can be 
integrated directly into the constraint network itself.

@@

The constraint network planner works on an evolutionary basis, similar to genetic algorithms. Starting from the goal 
(for example identifying a particular object in the environment), it will generate possible networks by recruiting 
existing primitive operations, linking them in, and checking whether the computation makes progress towards reaching the goal. 
The slots of a constraint are typed to limit what semantic operations will be tried next. Networks that perform useful 
functions are ``chunked'' into higher order units that then can be used as units in their own right, thus scaling up 
complexity. These chunked networks become associated with stereotyped grammatical expression, so that there is a very 
fast retrieval and semantic execution. \\

@@

1. The first strategy is the {\bfshape Basic Colour Strategy} which implements a prototype-based approach to 
the basic colour categories. This approach is similar to the one used in the Proper Naming Game, except that 
only the three colour-dimensions are profiled (the opponent channels red-green, yellow-blue and the brightness channel). 
The strategy is similar to the one shown earlier. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour-lenient ?s3 ?s2 ?cc)
(select-most-activated ?t ?s3)
\end{verbatim}
When the speaker says `green', this network is used together with a specific binding of ?cc to the prototype for 
green (denoted as [green]). 
\begin{verbatim}
(bind colour-category ?cc [green])
\end{verbatim}
A graphical representation of this network is shown in \figref{fig:basic-strat}.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/basic-strat.pdf}}
\caption{\label{fig:basic-strat} 
Constraint network for the Basic Colour strategy. The colours of each object in the context are retrieved, their 
similarity to the prototype computed, and the one with the highest similarity is chosen.}
\end{figure}

2. The second strategy is the {\bfshape Graded Membership Strategy} which categorises the topic first using 
a basic colour but then qualifies how well its colour fits with the prototype. For example, we can say `greenish' or
`very green'. Some languages (such as Tarahumara spoken in Northern Mexico) obligatory mark graded membership using 
suffixes. The core of the constraint network is defined as follows. A graphical representation of the network is shown
in \figref{fig:graded}. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour ?s3 ?s2 ?cc)
(filter-by-membership ?s4 ?s3 ?mc)
(select-most-activated ?t ?s4)
\end{verbatim}
This network grabs the objects in the context and binds them to ?s1 using the equal-to-context operation. 
The profile-colour-dimensions operation picks out the colour properties of the various objects in ?s1 resulting in 
a new set ?s2. The filter-by-colour operation computes for each object in the set ?s2
how well they match with the prototype bound to ?cc, constructing a new set ?s3. The filter-by-membership operation 
then computes how well 
each object in ?s3 satisfies the graded membership qualification bound to ?mc, 
constructing ?s4. And then the select-most-activated operation picks out the best candidate and binds it to ?t. 
Other data flows within the same network are possible, for example to find the bindings of ?mc and ?cc given 
a context and a referent ?t that needs to be identified.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/graded.pdf}}
\caption{\label{fig:graded} 
Constraint network implementing the Graded Membership Strategy. It builds further on the network for Basic Colour
Categorisation but adds a constraint to handle graded membership. }
\end{figure}

The network is entirely general for all expressions that use graded membership and gets instantiated for a particular 
case by specifying the semantic entities involved using the bind operation. For example, Tarahumara has the word 
`sita' for a particular colour prototype and `kame' for one type of graded membership: 
\begin{verbatim}
(bind colour-category ?cc [sita])
(bind membership-category ?mc [kame]) 
\end{verbatim}
